v--- \################################################################################ \## YOU CANNOT KNIT THIS FILE DIRECTLY FROM HERE! \## \## Instead use the wales-update-template.R script \## \################################################################################ output: bookdown::pdf_document2: toc: FALSE fontsize: 12pt header-includes: -

\usepackage{xcolor}

-   \usepackage{float}

    -   \usepackage{tabu}

    -   \usepackage{multirow}

    -   \usepackage{colortbl}

    -   \usepackage{flafter}

        params: current.year: 2017 dp.text: 1 dp.tables: 2 bibliography: wales.bib csl: apa-with-notes.csl ---

```{r setup, echo = FALSE, message= FALSE, warning=FALSE}
source(here::here("code/do-not-touch-scripts/wales.R"))
```

---
title: "`r paste("Local Authority Parking Finances in Wales", params$current.year, "-", params$current.year-1999)`"
nocite: |
  `r paste0(bib$refs, collapse = ", ")`
---

\renewcommand{\arraystretch}{1.2}

# Introduction

This note covers parking finances for the 22 local authorities in Wales. As in England and Scotland, local authorities are required to submit details of all their finances to the Welsh Government in a standard format. They are normally published in October, seven months after the financial year end. This report looks at the section on parking income and expenditure from `r FunFisc(4)` to `r FunFisc()` and is based primarily on data reported by Statistics for Wales (`r bib[bib$country == "Wales",]$refs`), as well as data reported by English and Scottish Local Government authorities; all of the sources are listed in the references.[^report-1]

[^report-1]: Contains public sector information licensed under the [Open Government Licence v3.0](http://www.nationalarchives.gov.uk/doc/open-government-licence/version/3/).

The published data is less comprehensive than in England and does not split out on- and off-street parking or show penalty income separately. The figures do not include any commercial off-street parking. In `r FunFisc()`, `r FunN2W(surplus.bin[2])` `r ifelse(surplus.bin[2] == 1, "council", "councils")` showed surpluses and `r FunN2W(surplus.bin[1])` showed deficits. All councils are now receiving income from parking.

# Summary

```{r summary, echo = FALSE, message = FALSE, warning = FALSE}
# print summary table
wal.summary.formatted %>% 
  kable( "latex", 
         booktabs = T, 
         digits = dp.tables, 
         longtable = TRUE,
         escape = F,
         align = c("r"),
         caption = "Summary of parking accounts for Wales (£ millions)",
         col.names = c("", "", yearz, paste0("Change ", current.year, "-", 
                                             current.year-1999, " on ", 
                                             current.year-1, "-", 
                                             current.year-1999-1))) %>% 
  collapse_rows(columns = 1, latex_hline = "major", valign = "middle") 

```

Table \@ref(tab:summary) shows the summary accounts for local authorities in Wales of the incomes and expenditures arising from parking charges and penalty income both for on- and off-street parking. The income has `r ifelse(wal.summary$change[1] < -0.05, paste("fallen by", FunDec(-wal.summary$change[1], dp.text), "%"), ifelse(wal.summary$change[1] > 0.05, paste("increased by", FunDec(abs(wal.summary$change[1]), dp.text), "%"),"stayed approximately the same"))`, the expenditure has `r ifelse(wal.summary$change[2] < -0.05, paste("fallen by", FunDec(-wal.summary$change[2], dp.text), "%"), ifelse(wal.summary$change[1] > 0.05, paste("increased by", FunDec(abs(wal.summary$change[2]), dp.text), "%"),"stayed approximately the same"))` and the surplus has `r ifelse(wal.summary$change[3] < -0.05, paste("fallen by", FunDec(-wal.summary$change[3], dp.text), "%"), ifelse(wal.summary$change[1] > 0.05, paste("increased by", FunDec(abs(wal.summary$change[3]), dp.text), "%"),"stayed approximately the same"))` compared to the previous fiscal year. Total transport expenditure has `r ifelse(wal.summary$change[4] < -0.05, paste("fallen by", FunDec(-wal.summary$change[4], dp.text), "%"), ifelse(wal.summary$change[1] > 0.05, paste("increased by", FunDec(abs(wal.summary$change[4]), dp.text), "%"),"stayed approximately the same"))` and the surplus now represents `r paste(FunDec(wal.summary[5, 6], dp.text), "%")` of net transport epxenditure. Since `r FunFisc(4)` income has `r ifelse(wal.summary[1,6] > wal.summary[1,2], "risen", "fallen")` by `r FunDec(abs(100*wal.summary[1,6]/wal.summary[1,2] - 100), dp.text)` % and expenditure has `r ifelse(wal.summary[2,6] > wal.summary[2,2], "risen", "fallen")` by `r FunDec(abs(100*wal.summary[2,6]/wal.summary[2,2] - 100), dp.text)` %. Over the same period the surplus has `r ifelse(wal.summary[3,6] > wal.summary[3,2], "risen", "fallen")` by `r FunDec(abs(100*wal.summary[3,6]/wal.summary[3,2] - 100), dp.text)` %. Figure \@ref(fig:fig1) gives a longer term overview of the trends in incomes, expenditures and surpluses.
